---
title: "101B_hw01_Daren_Sathasivam"
author: "Daren Sathasivam"
date: "2024-04-08"
output: pdf_document
---

# 1 In class activity: Making Meringues

Group's experiment design approach:

Group A(control): Using the ingredients - eggwhites, sugar, cream, and vanilla extract (general google recipe) \
Group B: Substituting eggwhites with aquafaba(liquid from canned chickpeas) - same portion as group A \
Group C: Substituting eggwhites with meringue powder - same portion as group A \

- Have several randomly assigned people make meringues based on listed recipe and repeat it several times.
- Measure weight/diameter of meringue
- Find the average size/weight for each group
- Compare results between each group

## a. What kind of approach do you use to design the experiment, best guess approach, one factor at a time approach, or factorial design approach?
- For the experiment, a factorial design approach is most suitable as it allows to test multiple different factors to understand how certain different recipes can change the size of the meringues. 

## b What factors do you consider to affect the outcome?
- Type and quantity of ingredients
- Mixing technique or method
- Baking temperature and duration
- Use of additional additives

## c. How will you measure the outcome variable?
- Weighing each meringue after baking to determine its weight
- Measuring the diameter of each meringue

## d. Which principle of experiment is used in your design? How do you implement the principles in your experiment?
- Randomization: Assigning participants to the different groups randomly helps ensure that any observed differences is not due to systematic biases(skilled vs unskilled bakers)
- Replication: Each recipe variation should be tested multiple times to obtain reliable and consistent results.

## e. Update the design of experiment based on what you hvae learned from Chapter 1. 
- Specifying the number or replications for each recipe to improve the precision of the estimates. 
- Control other confounding variables such as oven temperature, mixing technique, and no additives. 
- Blinding technique: participants do not know which substitute they are using to avoid any unconcsious bias when baking the meringue. 

# 2: Problem 2.25
```{r}
burn <- c(65, 82, 81, 67, 57, 59, 66, 75, 82, 70, 64, 56, 71, 69, 83, 74, 59, 82, 65, 79)
type <- c(rep(1, 10), rep(2, 10))
data2 <- data.frame(burn = burn, type = type)
# data2
```
## a. Test for $H_0: \sigma^2_1= \sigma^2_2$ vs. $H_a: \sigma^2_1 \neq \sigma^2_2$
```{r}
var.test(burn~type)
```
- After testing the hypothesis that the two variances are equal using $\alpha = 0.05$, the results yield a p-value of 0.9744, meaning we fail to reject the null hypothesis. This indicates that the variability in burn times does not significantly differ between the two types.  

## b. Test for $H_0: \mu_1= \mu_2$ vs. $H_a: \mu_1 \neq \mu_2$
```{r}
t.test(burn~type, var.equal = TRUE)
```
- The p-value for the t-test is 0.9622, meaning we fail to reject the null hypothesis. This indicates that the mean in burn times does not significantly differ between the two types. 

## c. Discuss the role of normality assumption in this problem. Check the assumption of normality for both types of flares.
```{r}
par(mfrow = c(1, 2))
qqnorm(burn[type==1]); qqline(burn[type==1])
qqnorm(burn[type==2]); qqline(burn[type==2])
```
- Both results seem to maintain the normality assumption. However, type1 flare deviates from the normal line relatively more in comparison to the type2 flare and also has a slight "S-shaped" curve which can indicate it containing heavier tails than a normal distribution. 

# 3: Problem 2.27 - skip(f) and (g)
```{r}
thickness <- c(11.176, 7.089, 8.097, 11.739, 11.291, 10.759, 6.467, 8.315, 5.263, 6.748, 7.461, 7.015, 8.133, 7.418, 3.772, 8.963)
temp <- c(rep(95, 8), rep(100, 8))
data3 <- data.frame(thickness = thickness, temp = temp)
# data3
```

## a. Is there evidence to support the claim that the higher baking temperature results in wafers with a lower mean photoresist thickness? Use $\alpha = 0.05$
```{r}
var.test(thickness~temp)
t.test(thickness~temp, var.equal = TRUE, alternative = "greater")
```
- The F-test's p-value is 0.5306 which is greater than $\alpha=0.05$. Thus, we fail to reject the null hypothesis of equal variances. Indicating that there is not enough evidence to conclude a significant difference in variances between the two temperature groups. The t-test with an alternative of "greater than" difference for the 95C group has a p-value of 0.009059. Thus we reject the null hypothesis indicating that there is statistically significant evidence to support the claim that higher baking temperatures results in wafers with a lower mean photoresist thickness. 

## b. What is the p-value for the test conducted in part (a)?
- F-test p-value: 0.5306
- t-test p-value: 0.009059

## c. Find a 95% confidence interval on the difference in means to provide a practical interpretation of this interval.
- Confidence Interval: $[0.8608158, \infty)$
- Interpretation: The confidence interval suggests that on average, wafers baked at 95C have 0.86 units greater than those baked at 100C. However, since the upper bound is $\infty$ and unboounded, it produces uncertainty and cannot definitively conclude how much higher the photoresist thickness may be in the 95C group.

## d. Draw dot diagrams to assist in interpreting the results from this experiment.
```{r}
data3$temp <- as.factor(data3$temp)
library(ggplot2)
p <- ggplot(data3, aes(x = temp, y = thickness, fill = temp)) +
  geom_dotplot(binaxis = 'y')
p + coord_flip()
```
- Based off of the dot diagrams, we cannot conclude for certain if the experiment follows a normal distribution. However, we can observe that the 100C group tends to result in a thinner photoresist layer relative to the 95C group. The model and observations can be improved by obtaining a larger sample size. 

## e. Check the assumption of normality of the photoresist thickness.
```{r}
par(mfrow = c(1, 2))
qqnorm(thickness[temp==95]); qqline(thickness[temp==95])
qqnorm(thickness[temp==100]); qqline(thickness[temp==100])
```

- Both temperature groups result in an "S-shaped" curve which indicates heavier tails than an expected normal distribution. A larger sample size may clarify this observation. 

# 4: Problem 2.30
```{r}
IQ <- c(6.08, 6.22, 7.99, 7.44, 6.48, 7.99, 6.32, 7.60, 6.03, 7.52, 5.73, 5.80, 8.42, 6.84, 6.43, 8.76, 6.32, 7.62, 6.59, 7.67)
order <- c(rep(1, 10), rep(2, 10))
data4 <- data.frame(IQ = IQ, order = order)
# data4
```
## a. Is the assumption that the difference in score is normally distributed reasonable?
```{r}
diff <- IQ[1:10]-IQ[11:20]
qqnorm(diff); qqline(diff)
```
- The normality assumption holds as the observed values on the QQ plot remain relatively near the line, suggesting that the difference in scores is approximately normally distributed. 


## b. Find a 95% confidence interval on the difference in mean score. Is there any evidence that mean score depends on birth order?
```{r}
t.test(IQ ~ order, paired = T)
```
- Confidence Interval: $[-0.366148, 0.2644148]$
- Since the confidence interval contains a 0, it indicates that there is no statistically significant difference in mean IQ scores between the two birth orders. 

## c. Test an appropriate set of hypotheses indicating that the mean score does not depend on birth order.
```{r}
t.test(IQ ~ order, paired = T)
```
- Null Hypothesis: $H_0: \mu_1 = \mu_2$
- Alternative Hypothesis: $H_1: \mu_1 \neq \mu_2$
- Based on the p-value of 0.723, we fail to reject the null hypothesis. This indicates that there is no statistically significant evidence to conclude a dependence of mean IQ scores on birth order in this experiment.  

